Chasing \rainbows{} with the Habitable Worlds Observatory

### 2023-03-16 src/tex/ms.tex
[showyourwork] overleaf sync

\subsection{Ocean Glint}
\cite{2001Natur.412..885F} noted the potential for time-resolved reflected light photometry to identify water oceans on the surface of an exoplanet.  \cite{2008Icar..195..927W} performed simulations of scattered light from directly imaged terrestrial exoplanets. 
%to demonstrate that phase-resolved photometry and polarization of a directly-imaged exoplanet would betray the presence of large surface oceans. 
They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases. %Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but
Specular reflection, a.k.a. glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., methane lakes on Titan;][]{2010GeoRL..37.7104S}. %\citet{lustig2019} explored mapping exoplanets at crescent phases using ocean glint.

In polarized light the glint appears at similar phase angles but its peak can be extended further particularly in the presence of atmospheric winds \citep{kopparla2018}.
%\cite{2006astro.ph.10518M} quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.  
\citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean.
%\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets. They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.

@@

\subsection{Polarized Phase Curves}
%compare to features in nonpol (strength rainbow); current capabilties?, errors from 3 papers,

@@

\subsection{Rainbows}
%calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below

@@

The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability for about 25 habitable zone planets.
%
This assertion requires an instrument with a coronograph capable of high contrast imaging in the optical-near infrared.
%
Following this release, NASA recently announced the start of the development of the Habitable Worlds Observatory (\hwo).
%
The precursor technology recommended by the Decadal Survey also lists ``direct imaging to probe polarized ocean glint on terrestrial planets'' as one of the priority capabilities \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
%
The performance and precise characteristics of this telescope (on-axis or off-axis, exact diameter, type of segmentation, type of coronagraph) are still to be determined.
%
The development will be heavily influenced by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
%
The Inner Working angle (\IWA) of the resulting telescope will be dependent on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.

@@

%\textbf{In this paper we...}
In this paper we quantify the accessible scattering phase angles for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds target stars. In \S 2.1 we describe the target list, in \S 2.2 we derive expressions for the range of scattering phase angles in a few different limits, in \S 2.3 we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4 we outline how the planet/star contrast changes as a function of orbital phase.  We present and discuss our results in \S 3 and conclude in \S 4.  %the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles  
% We minimally explore the contrast ratio as we are primarily exploring whether these scattering angles are reachable with inner working angle and outer working angle considered.

@@

\subsection{\hwo\ and Stellar Sample}
%\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
\hwo\ is still in the early stages of development, and the exact telescope and instrument design is undecided. 
%
In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
%
In addition, the observatory will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions. 
%
The \IWA\ of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations. 
%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory. 
%We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{-1}{\nano\meter}$ \todo{Update value!}.

>>>>

\subsection{Ocean Glint}
\cite{2001Natur.412..885F} noted the potential for time-resolved reflected light photometry to identify water oceans on the surface of an exoplanet.  \cite{2008Icar..195..927W} performed simulations of scattered light from directly imaged terrestrial exoplanets. 
%to demonstrate that phase-resolved photometry and polarization of a directly-imaged exoplanet would betray the presence of large surface oceans. 
They showed that oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases. %Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but
Specular reflection, a.k.a. glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., Titan's methane lakes;][]{2010GeoRL..37.7104S}. %\citet{lustig2019} explored mapping exoplanets at crescent phases using ocean glint.
Specular reflection appears at similar phase angles in both unpolarized and polarized light and its peak orbital phase can be extended by atmospheric winds, which produce waves and hence increase the size of the glint spot \citep{2011Icar..211..722B,kopparla2018}.
%\cite{2006astro.ph.10518M} quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.  
\citet{Takahashi2021} used polarimetric Earthshine measurements to detect the presence of the ocean, proving that the method works in practice for an unresolved world.
%\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets. They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.

@@

\subsection{Polarized Phase Curves}
\textcolor{blue}{Recap all of polarized phase curve work, including stuff not about ocean glint. Introduce the polarization signals of Rayleigh scattering, rainbow and glory.}
%compare to features in nonpol (strength rainbow); current capabilties?, errors from 3 papers,

@@

%\subsection{Rainbows}
%calc; retrieve species including for non terrestrials; droplet liklihood of liquid surface; lit review bailey et below

@@

\subsection{Habitable Worlds Observatory}
The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability for about 25 habitable zone planets.
%
This assertion requires an instrument with a coronograph capable of high contrast imaging in the optical-near infrared.
%
Following this release, NASA recently announced the start of the development of the Habitable Worlds Observatory (\hwo).
%
The precursor technology recommended by the Decadal Survey also lists ``direct imaging to probe polarized ocean glint on terrestrial planets'' as one of the priority capabilities \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
%
The performance and precise characteristics of this telescope (on-axis or off-axis, exact diameter, type of segmentation, type of coronagraph) are still to be determined.
%
The development will be heavily influenced by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
%
The Inner Working angle (\IWA) of the resulting telescope will be dependent on these choices, which in turn will have a significant impact on the expected exo-Earth yield \citep{Stark2019_exoplanetyield}.

@@

%\textbf{In this paper we...}
\subsection{Outline}
In this paper we quantify the accessible scattering phase angles for hypothetical terrestrial planets orbiting in the habitable zones of HabWorlds target stars. In \S 2.1 we describe the target list, in \S 2.2 we derive expressions for the range of scattering phase angles in a few different limits, in \S 2.3 we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets in these systems, and in \S 2.4 we outline how the planet/star contrast changes as a function of orbital phase.  We present and discuss our results in \S 3 and conclude in \S 4.  %the   the way that inclination and coronagraph inner working angle can limit the accessible phase angles  
% We minimally explore the contrast ratio as we are primarily exploring whether these scattering angles are reachable with inner working angle and outer working angle considered.

@@

\subsection{\hwo\ and Stellar Sample}
%\textcolor{blue}{Merge the 2 sections here into "Assumed technical specifications and sample"? I don't think this needs to be 2 subsections}
\hwo\ is still in the early stages of development, and the exact telescope and instrument design is undecided. 
%
In this work, we assume a \SI{6}{\meter} primary mirror, which is the currently favoured design for the observatory.
%
In addition, the observatory will utilise a coronagraph to suppress the stellar light and facilitate high contrast observations of their faint exoplanet companions. 
%
The \IWA\ of the coronagraph is currently undecided and will be driven by both the science requirements and technological limitations. 
%In this work, we investigate the scattering phase coverage obtainable for several choices for the \IWA\ of the observatory. 
%We also consider the Outer Working Angle (OWA) but for most targets this is not a limiting factor. 
%From the HabWorlds report. Inner Working Angle: The IWA defines the region near the star that cannot be accessed for direct imaging due to a coronagraphic mask, starshade obscuration, or an interferometric null. It depends on the architecture of the telescope, the starlight suppression system used, and the observation wavelength. While a nominal value could be defined by making assumptions about the design for Habitable Worlds Observatory, we choose instead to derive the IWA from the star list itself and Astro2020’s requirement that ~100 cumulative habitable zones be surveyed by direct imaging. As shown in the main table, an IWA near ~70 mas will be needed to access ~100 cumulative habitable zones, and would need to be achieved at all wavelengths of interest for spectral characterization. For any specific telescope and starlight suppression system architecture the IWA is typically proportional to wavelength; thus, the number of accessible HZs will strongly decrease as IWA(lambda) increases.

For ease of comparison, \cref{tab:IWA_OWA} indicates the angular separations corresponding to multiples of $\lambda / D$ for a \SI{6}{\meter} telescope at a wavelength of $\lambda = \SI{-1}{\nano\meter}$ \todo{Update value!}.
